# (PART) Cell culture {-}

# General protocols for cell culture

```{r setup_TC, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## General materials for cell culture:  

- culture medium (see cell protocols)  
- FCS  
- DMSO (D2650)  
- filters  
- formaldehyde (28906 or 28908)  

<br>
<br>

## Useful numbers

```{r TC_numbers, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, show_col_types=FALSE}

read_csv("cell_culture_numbers.csv") %>% knitr::kable()

```

<br>
<br>

## Yields:  

- podocyte progenitors (33deg): 12,500 cells per cm<sup>2</sup>  
- podocytes mature (37deg): 5,500 cells = 5 mcg protein per cm<sup>2</sup> 
- mCCD: 167,000 cells = 50 mcg protein per cm<sup>2</sup>  
- rpTECs: 320,000 cells per cm<sup>2</sup>  

Therefore confluent T75 will be ~10<sup>7</sup> mCCD cells.  

7 mcg RNA per 1 million cells.  

For mCCD cells: one confluent well of 6-well plate, lysed in 250 mcl RIPA gives c. 3000 mcg per ml = 750 mcg protein - i.e. plenty!  

<br>

## Protein

### Materials

Prefer to purchase RIPA buffer (ab156034) and PICs (Merck cocktail set III).  

<br>

**RIPA buffer (homemade)**:  

  - NaCl (5 M): 3 ml = 150 mM 
  - EDTA (0.5 M, pH 8.0): 400 mcl = 2 mM 
  - Triton-X 100: 1 ml = 1 % 
  - sodium deoxycholate: 1 g = 1 % 
  - Tris-HCl (1M, pH 8.0): 2.5 ml = 25 mM 
  - SDS (10 %): 1 ml = 0.1 % 
  - NaF: 210 mg = 50 mM 
  - Na3VO4 (200 mM): 0.5 ml = 1 mM 
  - Na β-glycerophosphate: 540 mg = 25 mM 
  - ddH2O adjust to 100 ml 
  - F.S. and store @ -20° as 5 ml aliquots
  
*Add protease inibitors (50 mcl of 100× to 5 ml; Merck cocktail set III) immediately prior to use.* 

<br>

**Protease inhibitors (100×) – cheap version**: 

  - NaN3: 21.5 mg = 330 mM 
  - AEBSF: 24 mg = 100 mM 
  - leupeptin: 200 mcl of 10 mM stock = 2 mM 
  - ddH2O to 1 ml 
  - store @ -20° as 65 mcl aliquots 

*For final (1 x) concentrations: NaN3 3.3 mM; AEBSF 1 mM; leupeptin 20 mcM. For urine / culture fluid (low protease activity) use at 1/10th of this – i.e. 20 mcl PIC for 20 ml fluid – NB need to compare this to others (e.g. Gonzales). NB 10 mM leupeptin stock = 1 mg in 210 mcl aq. Leupeptin 1 mg = £12.65; AEBSF 24 mg = £13 (therefore total £26 c.f. £45 for 1 ml Merck cocktail III).* 

<br>

### Method
  
**Preparation**:  

- pre-chill PBS and RIPA on ice 
- remember to add PIC to RIPA (e.g. Merck cocktail III @ 1:100) 

<br>

**Lysis on plate**:  

*volumes given here would be appropriate for 1x T75 flask of confluent cells*

- remove medium and wash cells twice with ice-cold PBS. 
- add 1.5mL ice-cold RIPA (with PIC). 
- incubate on ice × 15’, swirling occasionally. 
- gather lysate to one side using a cell scraper and transfer to microcentrifuge tube. 
- pass repeatedly (c. × 5) through a 21G (green) needle  
- spin @ 13000 g × 15’ @ 4 ° to pellet cell debris (= 12000 rpm on benchtop centrifuge in cold room)  

*(optional to increase yield – sonicate pellet for 30 s with a 50 % pulse)*

- transfer supernatant to a fresh tube
- store in 50 mcl aliquots @ - 80 °
- measure protein concentration (e.g. BCA assay in duplicate)

<br>

**Lysis after trypsinisation**:

- BEWARE - the trypsinisation may of course destroy epitopes  
- detach cells by trypsinisation
- wash thrice in 1.5mL ice-cold PBS
- resuspend pellet in 1.5mL RIPA (with PIC)
- pass repeatedly (c. × 5) through a 21G (green) needle
- incubate on ice × 15’ on rocking platform  
- then proceed as from spin step above   

<br>
  
## RNA

### Materials  

Trizol or Qiazol  

<br>

### Method

*For up to 10<sup>7</sup> cells, use 700 mcl QIAzol lysis reagent per reaction. This is supposed to be an appropriate volume for up to 10<sup>7</sup> cells or a dish up to 100 mm diameter. Old TRIzol protocols say 1mL for up to 10<sup>7</sup> cells.* 

- wash cells twice with PBS then detach by trypsinisation
- take aliquot to count cells if required

- transfer RNase-free tube; pellet at 160 g for 5 mins; completely aspirate spnt (= 1000 rpm on Heraus rotor 7591 on Biofuge Primo)

- flick tube to dislodge pellet
- add 700 mcl QIAzol lysis reagent
- homogenise cells:  
    - if processing < 3 x 10<sup>6</sup> cells, homogenize by vortexing for 1 min 
    - if processing > 3 x 10<sup>6</sup> cells, homogenize using syringe and needle 


- store homogenised cell lysate at – 70C

<br>

*Alternatives:*  
*- first potential alternative: freeze cell pellet*  
*- second potential alternative: lyse directly in the flask by adding 700 mcl QIAzol lysis reagent*  