# mCCD culture protocols


<!--
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```
--> 

## Materials  

- 500 ml DMEM/F-12 with L-glutamine or glutmax (e.g. Lonza 12-719F)  

- FBS (Gibco 10500)  

- tri-iodothyronine (Sigma T5516)  
    - *make a 50 mcM stock: dissolve 1mg in 1ml 1M NaOH then add 29 ml DMEM/F-12 and F.S.*   
    - *then dilute 1 ml in 49 ml DMEM/F-12 to give 1 mcM; store as 550 mcM aliquots*  
<br>

- dexamethasone (Sigma D8893)  
    - *add 510 mcl 100% EtOH to 1 mg dexamethasone to make 5 mM stock*  
<br>
    
- rhEGF (Sigma E9644)  
    - *make a 100 mcg/ml stock: dissolve 200 mcg in 2ml H2O and freeze as 55 mcl aliquots*  
<br>

- ITS (Gibco 41400045)  

- pen/strep (Gibco 15140122)  

- trypsin-EDTA (Gibco 253000054)  

<br>
<br>

## Methods

### Growth medium

To 500 ml DMEM/F12, add:  

- 5 ml ITS \*\*  
- 5 ml P/S \*\*  
- 500 mcl T3 \*\*  
- 5 mcl dexamethasone \*\*  
- 10 ml FCS	(don’t filter)  
- 50 mcl rhEGF (don’t filter)  

\*\* = combine & F.S.

Transfer 5 ml to T25 for a sterility test.  

<br>

```{r recipe, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, show_col_types=FALSE}

df_recipe <- read_csv("mCCD_growth.csv")
df_recipe %>% knitr::kable()

```

<br>
<br>

### Thawing

To thaw:  

- transfer 10 ml medium to a centrifuge tube and incubate @ 37deg    
- use 500 mcl aliquots of medium to thaw cells over 10 mins (i.e. squirt in 500 mcl, pipette up and down, transfer to 10 ml and repeat)  
- spin to pellet cells (1000 rpm for 5 mins = 290 *g*??), remove supernatant and re-suspend in 1 ml    
- add this to 9 ml in a T25    


### Splitting

To split:  

- remove medium and discard  
- wash x2 with PBS    
- incubate with trypsin-EDTA (e.g. 1 ml for T75) @ 37deg for 10 - 15 mins until detatched    
- inactivate with 10 vols medium (e.g. 10 ml for T75)  
- spin to pellet cells, discard supernatant and top up to desired volume with medium  
- split as desired (using haematocytometer to count cells if required)    


### Freezing

To freeze:  

- trypsinise cells and spin down  
- re-suspend in 10 % DMSO (filtered) in FCS and freeze as 1 ml aliquots in screw-top tubes    
- transfer to – 80deg in 'Mr. Frosty' chamber immediately and leave for 24 hrs    
- move to liquid nitrogen within a few days    


