# rpTEC-TERT1 & LLC-PK1 culture protocols

```{r setup_rpTEC, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## Materials (rpTEC-TERT1)

-	cell-culture grade water (Gibco 15230162)  

-	cell-culture grade NaOH (Sigma S2770)  

- 500 ml DMEM/F12 with L-glutamine (e.g. Gibco 11039021; used to use Biowhittaker BE12-719F and add L-glutamine separately)  
    - *NB if adding L-glutamine separately, then go for final concentration 2 mM*  

<br>

- tri-iodothyronine (Sigma T5516)  
    - *make a 100 ng/ml stock: dissolve 1mg in 1ml 1M NaOH, adjust to 50 ml with H2O (= 20 mcg/ml) and F.S.*  
    - *then dilute 50 mcl to 10 ml with H2O and freeze as 25 mcl aliquots*  
<br>

- hydrocortisone (Sigma H0135)  
    - *make a 50 mcg/ml stock: dissolve 1 mg in 1 ml EtOH, adjust to 20 ml with H2O and F.S.*  
    - *freeze as 370 mcl aliquots*  
<br>
    
- rhEGF (Sigma E9644)  
    - *make a 100 mcg/ml stock: dissolve 200 mcg in 2ml H2O and freeze as 55 mcl aliquots*  
<br>

- ITS (Gibco 41400045)  

- pen/strep (Gibco 15140122)  

<br>
<br>

## Methods (rpTEC-TERT1)  

```{block2, type='note'}
This growth medium adapted from various sources, including the ATCC documentation, Aschauer et al. (Mol Cell Biol 2013), Sneddon et al. (JBC 2016), Wieser et al. (AJP 2008).  

NB ATCC documentation states that RPTEC/TERT1 cells should be grown in G418 (100 mcg / ml) in order to maintain selective pressure for immortalisation.  This does not appear to be important in practice.  It is not entirely clear why rpTEC-TERT1 cells are cultured in serum-free medium.  This is perhaps because of historic serum-free protocols used to culture rTECs in the 1970s (Wieser et al., 2008).  For the best discussion of this need to read Taub & Livingston (1981).  

ITS was classically used to allow cells to grow in serum-free or low-serum (2-4 %) media.  There are some differences in commercial formulations.  For example, Invitrogen 41400 (diluted 1:100): I = 10 mg/L, T = 5.5 mg/L, S = 6.7 mcg/L; Sigma 1884 (used as per Bristol protocol for rpTECs): I = 5 mg/L, T = 5 mg/L, S  = 5 mcg/L; Sigma 3146 (diluted 1:100): I = 10 mg/L, T = 5.5 mg/L, S = 5 mcg/L.  

```

See also [ATCC protocols](https://www.atcc.org/products/crl-4031).  


### Growth medium

To 500 ml DMEM/F12, add:  

- 5 ml L-glutamine (ONLY IF NOT ALREADY IN BASE MEDIUM) \*\*  
- 5 ml ITS \*\*  
- 5 ml P/S \*\*  
- 20 mcl T3  
- 360 mcl hydrocortisone  
- 50 mcl rhEGF (don’t filter)  

\*\* = combine & F.S.

Transfer 5 ml to T25 for a sterility test.  

NB - this is serum-free medium.  

<br>

```{r recipe_rpTEC, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, show_col_types=FALSE}

read_csv("rpTEC_growth.csv") %>% knitr::kable()

```

<br>

### Passaging

Typically need to trypsinise for c. 7 -- 8 mins.  Remember to quench with a serum-containing medium.  Typically split when 90% confluent, at ratios 1:4 -- 1:6.  Cells look unhappy for 24 -- 48 hrs after splitting.  

Freeze in 10% DMSO / 90% FCS (or 10% DMSO / 90% culture medium).  

<br>
<br>

## LLC-PK1 cells  

See [ATCC protocols](https://www.atcc.org/products/cl-101).  

Grow in Medium 199, supplemented with 3% FCS and P/S.  Typically split when 80% confluent at a ratio of 1:3 -- 1:8.  If they become too confluent, then hard to trypsinise.  Freeze in 5% DMSO / 95% growth medium.  

<br>
<br>